---
title: "visual-phrase-structure"
---

# Overview

Here are our notes:

-   [Summary of experiments we will publish](https://georgetown.box.com/s/nanuy5nxxuydx9vmlo06igssq9ht2nkj) (note: does not yet include the additional experiments that RA suggested including on Friday 6/5/26)
-   [Full summary of all experiments](https://docs.google.com/document/d/162QLIBBQVdeLherblKcRbrBdSM18KBoe/edit#heading=h.mvxyavvrd69v)

```{r, include=FALSE}
library(tidyverse)
library(gt)
library(knitr)

# set up variables ----------------------
names_to_keep <- c("Flying shapes, 1-shape window",
                "Flying shapes, 2-shape window",
                "Looming shapes",
                "Base pairs, static, complete grid",
                "Base pairs, static, limited grid",
                "Base pairs, phrase-by-phrase")

condition_order <- c("Phrase-structure","Control")

factor_variables <- c("version","participant", "participants","session", "condition", "test", "paradigm", "nback", "materials", "test", "name", "experiment")

learning_cutoff <-  .75  # cutoff for defining "good learners"

# load files ----------------------
version_info <- read.csv("data/tn_visual_version_summary.csv") %>%
    select(version:experiment)

individual_means_raw <- read.csv("data/tn_visual_all_data.csv") %>%
  merge(version_info, by="version")


```

```{r, include=FALSE}

# tidy data ----------------------
data <- individual_means_raw %>%
  filter(name %in% names_to_keep)%>%
  filter(version !="v5")%>% # wrong data, need to replace
  mutate(condition=case_match(condition,
                              "ctl" ~ "Control",
                              "control" ~ "Control",
                              "exp" ~ "Phrase-structure",
                              .default = condition))%>%
  rename(indiv_mean=indivMean)%>%
  arrange(version,participant)%>%
#  select(-version)%>%
#  rename(version=version_combined)%>%
#  unique()%>%   # figure out why there are 40 extra rows. Data loaded twice???
  mutate(across(all_of(factor_variables), as.factor))%>%
  mutate(condition=factor(condition, condition_order, ordered = TRUE))%>%
  mutate(name=factor(name, names_to_keep, ordered = TRUE))%>%
  droplevels()

# clean data  ----------------------

# some people apparently participated in more than one experiment
data %>% select(participant, name, version)%>%unique()%>%
  group_by(participant) %>% 
  summarise(experiments = paste(version, collapse = ", "),
            n_experiments=n_distinct(version)) %>%
  arrange(desc(n_experiments))

# remove data for all but the first experiment they participated in
data <- data %>% 
  filter(!(participant=="5ebe172a6d155d018177a84d" & version == "v12")) %>%
  filter(!(participant=="60b1a6ceb8ed805efaa99650" & version == "v18"))

# confirm that now each participant is only in there once
data %>% select(participant, name, version)%>%unique()%>%
  group_by(participant) %>% 
  summarise(experiments = paste(version, collapse = ", "),
            n_experiments=n_distinct(version)) %>%
  arrange(desc(n_experiments))

summary(data)

```

## Ns by experiment

Here are the experiments we're going to publish, with Ns. A few Ns are different. from what's in our summary sheets because I discovered that a couple of participants were run in multiple experiments. I removed data for all but the first experiment they participated in.

```{r}
data %>% select(experiment, name, version_combined, condition, participant) %>%
  unique()%>%
  group_by(experiment, name, condition)%>%
  summarise(
    versions = paste(sort(unique(version_combined)), collapse = ", "),
    n = n_distinct(participant),
    .groups = "drop")%>%
  pivot_wider(
    names_from = condition, 
    values_from = n,
    values_fill = 0)%>%
  arrange(experiment) %>%
  kable()


# SOMEHOW, data for v21 (controls) got added twice, but only for some participants. i went directly into the data file 6/9/26 and deleted the extras, i think. 

```

# Results: All learners

```{r}
group_means <- data %>% 
  group_by(experiment, name, condition, session, test)%>%
  summarise(
    group_mean = mean(indiv_mean), 
    n = n(),
    sd = sd(indiv_mean),
    .groups = "drop")%>%
  mutate(
    se = sd/sqrt(n),
    across(c(group_mean, sd, se), ~round(.x, 2)))%>%
  arrange(name, condition, session, test)

group_means_by_experiment <- data %>% 
  group_by(experiment, condition, session, test)%>%
  summarise(
    group_mean = mean(indiv_mean), 
    n = n(), 
    sd =sd(indiv_mean),
    .groups = "drop")%>%
  mutate(
    se =sd/sqrt(n),
    across(c(group_mean, sd, se), ~round(.x, 2)))%>%
  arrange(experiment, condition, session, test)
```

## Phrase Test

### By version

```{r}
group_means %>% 
  filter(test=="phrase")%>%
  select(-test)%>%
  kable()


group_plot <- ggplot(
    filter(group_means, session=="2" & test=="phrase"), 
    aes(x = condition, y = group_mean))+
  geom_bar(
    stat = "identity", 
    aes(fill = condition), 
    color = "black",
    position=position_dodge(preserve="single"))+
  geom_errorbar(
    aes(ymin = group_mean - se, ymax = group_mean + se), 
    width = .25, 
    position=position_dodge(width=.9, preserve="single"))+
  geom_hline(yintercept = .5, linetype = "dashed", alpha = .5)+
  ylim(0,1)+
  scale_fill_brewer(palette="Set1")+
  facet_wrap(~name, labeller = label_wrap_gen(width = 20))+
  theme_bw()

#group_plot

group_plot +
  geom_violin(
    data = filter(data, session =="2" & test == "phrase"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="phrase"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    binaxis="y", 
    stackdir="center")+
  labs(title="Phrase Test (Day 2, all participants)")


```

### By experiment

```{r}
group_means_by_experiment %>%
  filter(test=="phrase")%>%
  select(-test)%>%
  kable()

group_plot %+%
  filter(group_means_by_experiment, session=="2" & test=="phrase")+
  facet_wrap(~experiment, labeller = label_wrap_gen(width = 20))+
  geom_violin(
    data = filter(data, session =="2" & test == "phrase"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="phrase"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    dotsize = .5,
    binaxis="y", 
    stackdir="center")+
  labs(title="Phrase Test (Day 2, all participants)")
```

## Sentence Test

### By version

```{r}
group_means %>% 
  filter(test=="sentence")%>%
  select(-test)%>%
  kable()

group_plot %+% 
  filter(group_means, session=="2" & test=="sentence")+
  geom_violin(
    data = filter(data, session =="2" & test == "sentence"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+  
  geom_dotplot(
    data = filter(data, session=="2" & test=="sentence"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    binaxis="y", 
    stackdir="center")+
  labs(title="Sentence Test (Day 2, all participants)")
```

### By experiment

```{r}
group_means_by_experiment %>%
  filter(test=="sentence")%>%
  select(-test)%>%
  kable()

group_plot %+%
  filter(group_means_by_experiment, session=="2" & test=="sentence")+
  facet_wrap(~experiment, labeller = label_wrap_gen(width = 20))+
  geom_violin(
    data = filter(data, session =="2" & test == "sentence"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="sentence"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    dotsize = .5,
    binaxis="y", 
    stackdir="center")+
  labs(title="Sentence Test (Day 2, all participants)")
```

# Results: Good learners

### Defining "good learners"

Throughout the project, we've looked at results for everyone combined and also only for "good learners", defined as people who surpassed a certain threshold on the phrase test. We thought this was necessary because we didn't always have an attention check built into the paradigm.

Here, I've defined the threshold for "good learners" as a score of **≥ `r learning_cutoff`** on the phrase test.

```{r, echo=FALSE}
good_learners <- data %>% spread(session,indiv_mean)%>%
  rename("session_1" = "1", "session_2"="2")%>%
  # update this for v4 since it had 3 days
  mutate(session_last = ifelse(is.na(session_2),session_1,session_2))%>%
  filter(test=="phrase")%>%
  mutate(learning = ifelse(session_last >= learning_cutoff, "good","not_good"))%>%
  filter(learning=="good")%>%
  droplevels()%>%
  .$participant

data$phrase_learning = ifelse(data$participant %in% good_learners,"good","not_good")

data %>% 
  select(name, version_combined, phrase_learning, condition, participant) %>%
  unique() %>%
  group_by(name, condition, phrase_learning) %>%
  summarise(
    versions = paste(sort(unique(version_combined)), collapse = ", "),
    n = n_distinct(participant),
    .groups = "drop"
    ) %>%
  pivot_wider(
    names_from = phrase_learning,
    values_from = n,
    values_fill = 0)%>%
  mutate(good_proportion = good/(good+not_good))%>%
  arrange(name, versions, condition)%>%
  kable()
```

```{r}
group_means_by_learner_performance <- data %>% 
  group_by(name, condition, session, test, phrase_learning)%>%
  summarise(
    group_mean = mean(indiv_mean), 
    n = n(),
    sd = sd(indiv_mean))%>%
  mutate(
    se = sd/sqrt(n),
    across(c(group_mean, sd, se), ~round(.x, 2)))%>%
  arrange(name, condition, session, test)


group_means_by_experiment_by_learner_performance <-  data %>% 
  group_by(experiment, condition, session, test, phrase_learning)%>%
  summarise(
    group_mean = mean(indiv_mean), 
    n = n(),
    sd = sd(indiv_mean))%>%
  mutate(
    se = sd/sqrt(n),
    across(c(group_mean, sd, se), ~round(.x, 2)))%>%
  arrange(experiment, condition, session, test)
```

## Phrase Test

Phrase test performance is what we filtered by, so of course it's good. I'm just including these results for completeness.

### By version

```{r}
group_means_by_learner_performance %>%
  filter(test=="phrase")%>%
  select(-test)%>%
  kable()

group_plot %+%
  filter(group_means_by_learner_performance, session=="2" & test=="phrase" & phrase_learning=="good")+
  geom_violin(
    data = filter(data, session =="2" & test == "phrase" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="phrase" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    binaxis="y", 
    stackdir="center")+
  labs(title="Phrase Test (Day 2, 'good learners' only)")

```

### By experiment

```{r}
group_means_by_experiment_by_learner_performance %>%
  filter(test=="phrase")%>%
  select(-test)%>%
  kable()

group_plot %+%
  filter(group_means_by_experiment_by_learner_performance, session=="2" & test=="phrase" & phrase_learning=="good")+
  facet_wrap(~experiment, labeller = label_wrap_gen(width = 20))+
  geom_violin(
    data = filter(data, session =="2" & test == "phrase" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="phrase" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    dotsize=.5,
    binaxis="y", 
    stackdir="center")+
  labs(title="Phrase Test (Day 2, 'good learners' only)")
```

## Sentence Test

```{r}
group_means_by_learner_performance %>%
  filter(test=="sentence")%>%
  select(-test)%>%
  kable()


group_plot %+%
  filter(group_means_by_learner_performance, session=="2" & test=="sentence" & phrase_learning=="good")+
  geom_violin(
    data = filter(data, session =="2" & test == "sentence" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="sentence" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    binaxis="y", 
    stackdir="center")+
  labs(title="Sentence Test (Day 2, 'good learners' only)")
```

```{r}
group_means_by_experiment_by_learner_performance %>%
  filter(test=="sentence")%>%
  select(-test)%>%
  kable()

group_plot %+%
  filter(group_means_by_experiment_by_learner_performance, session=="2" & test=="sentence" & phrase_learning=="good")+
  facet_wrap(~experiment, labeller = label_wrap_gen(width = 20))+
  geom_violin(
    data = filter(data, session =="2" & test == "sentence" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition),
    alpha = .2,
    color = "grey"
  )+
  geom_dotplot(
    data = filter(data, session=="2" & test=="sentence" & phrase_learning=="good"),
    aes(x = condition, y = indiv_mean, fill = condition), 
    alpha = .2, 
    dotsize=.5,
    binaxis="y", 
    stackdir="center")+
  labs(title="Sentence Test (Day 2, 'good learners' only)")
```
